Challenges in Developing Great Quasi-Monte Carlo Software

### 2023-06-18 MCQMC2022ChallengesQMCSoftwareArticle_revision_1.tex
SC: change f to g to make it clearer

We illustrate in Figure~\ref{CDHJSfigbeam} how UM-Bridge connects QMCPy with a differential equation boundary value problem solver for a cantilevered beam, with a random Young's modulus \SCComment{$m(x)$ of length $L$, assumed to be piecewise constant over three regions, each of which with a uniform distribution of uncertainty~\cite{ParSee22a}, i.e., $m(x) = \left[\begin{matrix} m_1(x), m_2(x), m_3(x) \end{matrix} \right]$ with $m_i(x) \sim \calU[1,1.2]^3$.  The beam position $x$ is discretized at $N=31$ equidistant points, and 
the vertical deflection $u(x)$ when the beam is subject to a given external load  $g(x)$ satisfies the Euler-Bernoulli partial differential equation (PDE):
$$ \frac{\partial^2}{\partial x^2}\left[ \exp(m(x)) \frac{\partial^2  u(x)}{\partial x^2}\right] = g(x),$$
%where $m(x) = \log E(x)$ is the log of an effective stiffness $E(x)$.  The  
with the following boundary conditions:
$$ \left.u(x)\right|_{x=0} = 0, \qquad 
\left.\frac{\partial u(x)}{\partial x}\right|_{x=0} = 0, \qquad
\left.\frac{\partial^2 u(x)}{\partial x^2}\right|_{x=L} = 0, \qquad  \left.\frac{\partial^3 u(x)}{\partial x^3}\right|_{x=L} = 0. $$
This PDE can be discretized with finite differences yielding a linear system $K(m) \hat{u} = \hat{f}$, where $\hat{u}$ and $\hat{f} \in \mathbb{R}^N$. }
%approximates $u(x)$ and $m(x)$ at the nodes, respectively. 
%The forward model $F:\mathbb{R}^3 \rightarrow \mathbb{R}^N $ maps $m = \left[\begin{matrix}m_1, m_2, m_3 \end{matrix}\right]^T$ to $\hat{u}$, while $Q : \mathbb{R}^N \rightarrow \mathbb{R}^2 $ selects the estimated solution at the 10th and 25th nodes, i.e., $ Q(\hat{u}) := \left[\begin{matrix} \hat{u}_{10}, \hat{u}_{25} \end{matrix} \right]^T$. The objective is to determine the distributions of $Q(F(m))$. 
QMCPy simulates the random Young's modulus and computes the expected deflection of the beam at the node points for a sequence of error tolerances using lattice rules.  By taking advantage of virtual machines and docker containers, UM-Bridge provides automatic load balancing to facilitate seamless parallel model evaluation.% the evaluation of the beam deflection can be computed in parallel in multiple cores of a single CPU to decrease execution time.

>>>>

We illustrate in Figure~\ref{CDHJSfigbeam} how UM-Bridge connects QMCPy with a differential equation boundary value problem solver for a cantilevered beam, with a random Young's modulus \SCComment{$m(x)$ of length $L$, assumed to be piecewise constant over three regions, each of which with a uniform distribution of uncertainty~\cite{ParSee22a}, i.e., $m(x) = \left[\begin{matrix} m_1(x), m_2(x), m_3(x) \end{matrix} \right]$ with $m_i(x) \sim \calU[1,1.2]^3$.  The beam position $x$ is discretized at $N=31$ equidistant points, and 
the vertical deflection $u(x)$ when the beam is subject to a given external load  $g(x)$ satisfies the Euler-Bernoulli partial differential equation (PDE):
$$ \frac{\partial^2}{\partial x^2}\left[ \exp(m(x)) \frac{\partial^2  u(x)}{\partial x^2}\right] = g(x),$$
%where $m(x) = \log E(x)$ is the log of an effective stiffness $E(x)$.  The  
with the following boundary conditions:
$$ \left.u(x)\right|_{x=0} = 0, \qquad 
\left.\frac{\partial u(x)}{\partial x}\right|_{x=0} = 0, \qquad
\left.\frac{\partial^2 u(x)}{\partial x^2}\right|_{x=L} = 0, \qquad  \left.\frac{\partial^3 u(x)}{\partial x^3}\right|_{x=L} = 0. $$
This PDE can be discretized with finite differences yielding a linear system $K(m) \hat{u} = \hat{g}$, where $\hat{u}$ and $\hat{g} \in \mathbb{R}^N$. }
%approximates $u(x)$ and $m(x)$ at the nodes, respectively. 
%The forward model $F:\mathbb{R}^3 \rightarrow \mathbb{R}^N $ maps $m = \left[\begin{matrix}m_1, m_2, m_3 \end{matrix}\right]^T$ to $\hat{u}$, while $Q : \mathbb{R}^N \rightarrow \mathbb{R}^2 $ selects the estimated solution at the 10th and 25th nodes, i.e., $ Q(\hat{u}) := \left[\begin{matrix} \hat{u}_{10}, \hat{u}_{25} \end{matrix} \right]^T$. The objective is to determine the distributions of $Q(F(m))$. 
QMCPy simulates the random Young's modulus and computes the expected deflection of the beam at the node points for a sequence of error tolerances using lattice rules.  By taking advantage of virtual machines and docker containers, UM-Bridge provides automatic load balancing to facilitate seamless parallel model evaluation.% the evaluation of the beam deflection can be computed in parallel in multiple cores of a single CPU to decrease execution time.
